CrypTool Script\footnote{ Background reading for the CrypTool program}\\ Mathematics and Cryptography

### 2003-04-22 script/en/chapters/introduction.tex
be updates

In order to also enable developers/authors to work together independently 
the topics have been split up and for each topic an extra chapter has been 
written which can be read on its own. The later editorial work in TeX added 
cross linkages between different sections and footnotes describing where you
can find the according 
functions within the CrypTool\index{CrypTool} program. Naturally 
there are much more interesting topics in mathematics and cryptography 
which could be discussed in greater depth -- therefore this is only one of many
ways to do it.

@@

I am deeply grateful to all the people helping with their impressive
commitment who have made this global project so successful. Especially I would
like to acknowledge the English language proof-reading of this script version
done by Richard Chistenson and L. Montgomery.

@@

The {\em authors} have attempted to describe cryptography for a broad 
audience -- without being mathematically incorrect. We believe that this
didactical pretension is the best way to promote the awareness for IT
security and the readiness to use standardised modern cryptography.
\par \vskip + 15pt

>>>>

In order to also enable developers/authors to work together independently 
the topics have been split up and for each topic an extra chapter has been 
written which can be read on its own. The later editorial work in TeX added 
cross linkages between different sections and footnotes describing where you
can find the according functions within the CrypTool\index{CrypTool} 
program (see menu tree in appendix A). 
% \hyperlink{appendix-menutree}{(see menu tree in appendix A).}
% in \ref{s:appendix-menutree}
% \hypertarget{appendix-menutree}{}\label{s:appendix-menutree}
Naturally there are much more interesting topics in mathematics and
cryptography which could be discussed in greater depth -- therefore this
is only one of many ways to do it.

@@

I am deeply grateful to all the people helping with their impressive
commitment who have made this global project so successful. Especially I would
like to acknowledge the English language proof-reading of this script version
done by Richard Christenson and L. Montgomery.

@@

The authors
% \hyperlink{appendix-authors}{(authors)}
% in \ref{s:appendix-authors}
% \hypertarget{appendix-authors}{}\label{s:appendix-authors}
have attempted to describe cryptography for a broad 
audience -- without being mathematically incorrect. We believe that this
didactical pretension is the best way to promote the awareness for IT
security and the readiness to use standardised modern cryptography.
\par \vskip + 15pt

### 2003-04-22 script/en/chapters/moderncryptography.tex
be updates

In this process, we describe a problem as ``easy'' if it can be solved in 
\index{Runtime!polynomial} polynomial time as a function of the length of the 
input. If the length of the input is $ n $ bits, then the time for calculating 
the function is proportional to $ n^{a}, $ where $ a $  is a constant. We say 
that the \index{Complexity} complexity of such problems is $ O(n^{a}) $.

The term ``practically impossible'' is slightly less precise. In 
general, we can say that a problem cannot be solved \index{Runtime!efficient} 
efficiently, if the time required to solve it increases more quickly than the 
polynomial time as a function of the size of the input. If, for example, the 
length of the input is $ n $  bits and the time required for calculating the 
function is proportional to $ 2^n $, then the following currently applies: the 
function practically cannot be calculated for $n > 80$ (for $ a=5 $ the 
following applies: from the length $ n=23 $, $ 2^n$ is greater than $n^5 $, after 
which $ 2^n $ clearly increases more quickly ($ 2^{23}=8,388,608, $ $ 23^5= 
6,436,343 $)).\par

@@

\vskip +10 pt
{\bf RSA procedure} (based on the factorisation problem).
\vskip +20 pt
\index{Factorisation problem}
\index{Euler!(phi) function}
{\bf Comment:} 
The Euler phi function is defined as: $ \phi (N)$ is the number of natural 
numbers that do not have a common factor with $ N $ \index{Co-prime} {}
$ x \leq N. $ Two natural numbers $ a $ and $ b $ are co-prime if 
$ \ggt (a,b)=1. $

>>>>

In this process, we describe a problem as ``easy'' if it can be solved in 
\index{Runtime!polynomial} polynomial time as a function of the length of the 
input. If the length of the input is $ n $ bits, then the time for calculating 
the function is proportional to $ n^{a}, $ where $ a $  is a constant. We say 
that the \index{Complexity} complexity of such problems is $ O(n^{a}) $
t (Landau- or Big-O notation).

If you compare 2 functions  $ 2^n $  and   $ n^{a} $,  where $ a $  is a
constant, then there always exists a value for  $ n $,  from which for all
further  $ n $  applies: $ n^{a}  <  2^n $. 
The function  $ n^{a} $  has a lower complexity.    
Sample: for $ a=5 $ the following applies: from the length $ n=23 $, $ 2^n$
is greater than $n^5 $; for further $ n $  $ 2^n $ clearly increases more
quickly \
[($ 2^{22}= 4,194,304 $, $ 22^5= 5,153,632 $); \
 ($ 2^{23}= 8,388,608 $, $ 23^5= 6,436,343 $); \
 ($ 2^{24}=16,777,216 $, $ 24^5= 7,962,624 $);].\par

The term ``practically impossible'' is slightly less precise. In 
general, we can say that a problem cannot be solved \index{Runtime!efficient} 
efficiently, if the time required to solve it increases more quickly than the 
polynomial time as a function of the size of the input. If, for example, the 
length of the input is $ n $  bits and the time required for calculating the 
function is proportional to $ 2^n $, then the following currently applies: the 
function practically cannot be calculated for $n > 80$.

@@

\vskip +10 pt
{\bf RSA procedure} (based on the factorisation problem).
\vskip +20 pt
\index{Factorisation!factorisation problem}
\index{Euler!(phi) function}
{\bf Comment:} 
The Euler phi function is defined as: $ \phi (N)$ is the number of natural 
numbers that do not have a common factor with $ N $ \index{Co-prime} {}
$ x \leq N. $ Two natural numbers $ a $ and $ b $ are co-prime if 
$ \ggt (a,b)=1. $
